### 2014-12-04 docs/ms.Rmd
Continued shaping intro, drafted half of data/code results section

# Reproducible research in the study of biological colouration

@@

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field []. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent developments in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. While such advances are positive for the field, the rapid rate of progression means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported (Fig. 1).

Replication and transparency lie at the heart of the scientific enterprise. The full replication of individual studies based on independently collected data is the gold standard in cumulative science. This may often be unfeasible however, especially in field-based research, which includes a great deal of research on colour-traits []. In contrast, reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility is of tremendous value both to individual researchers and the broader community as it ensures greater comparability between studies, and may provide a launching point for future research.

Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline simple ways in which they may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration. We then present a set of guidelines for the reporting of methods, and we assess a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also analyse the availability of publicly archived data and code, and suggest some tools that may be useful for increasing reproducibility.

@@

_The curent state of methods reporting in the literature_

@@

While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods. This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria. Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of data collection and analysis methods within such studies.

@@

todo: results here

_Tools for reproducible research_

Data - archiving (figshare, dryad), raw data preferable to processed data, concerns over data sharing, embargos are fine just get it archived.
Code - benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

@@

>>>>

# Reproducible research in the study of colour in nature

@@

Studies of biological colouration have generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The resurgent interest in biological colouration has in large part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [@endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. These developments are positive for the field, and have allowed increasing numbers of researchers to rapidly collect and analyse ever greater amounts of data. The rapid rate of progression, however, means that a deep understanding of common methods is increasingly beyond the grasp of the average researcher. As a consequence, the subtle complexity of many analytical techniques [e.g. the 3-7 parameters that define a typical visual model; @endler_comparing_2005; @vorobyev_receptor_1998] is often overlooked by empiricists, which leads to critical methodological information often not being reported.

Replication and transparency lie at the heart of the scientific enterprise. The full replication of individual studies may often be infeasible @peng_reproducible_2011, but reproducibility - the re-analysis of a study's data by independent researchers - has been presented as a readily attainable alternative [@peng_reproducible_2006; @peng_reproducible_2011]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly archived and available, and (iii) analysis code is made publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the simple practice of complete methods reporting and public data archiving is of tremendous value. It ensures greater comparability between studies, allows for the the independent verification of results, and provides a foundation for the testing of new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007].

Our aim here to explore current levels of reproducibility in studies of biological colouration, and to outline ways in which they may be improved in line with points (i) - (iii) above. We begin by briefly outlining some of the most common methods for studying biological colouration. We then present a set of guidelines for the comprehensive reporting of methods, and we analyse a subset of the literature against these criteria to assess the current state of methods reporting in colour science. We also assess the availability of publicly archived data and code, and suggest some tools that may be useful for increasing the reproducibility of colour trait research.

@@

_The current state of methods reporting in the literature_

@@

While some of the percentages reported above seem alarming, it is important to note that around one-third of papers (`r round(ref_stud, digits = 2)`%) made reference to previous work for various methods. This means that our above figures are slightly liberal given that the referenced works may have comprehensively covered some of these criteria. Despite this, two-thirds of the studies did not reference previous work, and were missing important methodological details, suggesting the need for more comprehensive reporting of methods.

@@

Of the `r nrow(coldat)` studies analysed, (PERCENTAGE)% publicly provided the raw underlying data, (PERCENTAGE)% provided data in a pre-processed form, and (PERCENTAGE)% of studies did not provide any publicly accessible data. The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@vines_mandated_2013; @wolkovich_advances_2012], though this trend is gradually shifting as funders and journals increasingly mandate the release of data [@whitlock_data_2010]. The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods. This is particularly relevant to the field of colour-science, where new methods for analysing colour data are regularly developed [e.g. @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014]. The provision of open data may also foster collaborations as researchers draw upon existing work, and increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007]. Given the increasing rate of retractions [@steen_retractions_2010; @van_trouble_2011], there is also impetus for researchers to maintain credibility through transparency. We thus encourage researchers to publicly archive data in as raw a form as possible (e.g. individual reflectance spectra). Several excellent data repositories now exist, such as figshare and dryad, and the preparation of data is relatively straightforward [@whitlock_data_2011]. The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data, and are subsequently credited [@neylon_science_2012].

None of the included studies linked to any form of code...benefits of code over point-and-click, open source shit (R, pavo, things), github, sourcefourge etc.

@@

The permanent private storage of data [or, more often, the complete loss of data; @alsheikh_public_2011] is increasingly difficult to justify, and the benefits of open data are considerable.
